Regular singular Volterra equations on complex domains

### 2023-08-15 reg-sing-volterra.tex
Finish revising proof that prototype solution works

\item Replace $\epsilon$ with $\gamma$.

@@

Our definition of a regular singularity implies that as $|\zeta|$ goes to zero, $|p|$ also goes to zero, while $|q|$ settles into a compact interval that doesn't include zero.
\begin{verify}
[\textit{Proof.} Fix any $a'$ in the neighborhood where the difference above is bounded. By confining $a$ to small enough neighborhoods of $\zeta = 0$, we can make $|\tau/\zeta(a)|$ arbitrarily large, forcing $|q(a')/p(a)|$ to become arbitrarily large to satisfy the definition. Since $a'$ is fixed, the only way for this ratio to get arbitrarily large is for $|p(a)|$ to get arbitrarily small.]
\end{verify}
For some of our results, we'll also need to control $|p|$ and $|q|$ in various ways as $|\zeta|$ goes to infinity.
\begin{conditions}
\condconst{slow}{\lambda_0}\label{cond:slow} We can find constants $\lambda_p$ and $\lambda_\text{ratio}$ with
\begin{align*}
\left|\frac{1}{p}\right| & \lesssim e^{\lambda_p |\zeta|} &
\left|\frac{q}{p}\right| & \le \lambda_\text{ratio}
\end{align*}
outside some neighborhood of $\zeta = 0$ in $\Omega$. Note that this only constrains the ratio of $p$ and $q$ when they're evaluated at the same point. The constants will only enter our reasoning through their sum, $\lambda_0 = \lambda_p + \lambda_\text{ratio}$.
\end{conditions}
This condition explains why $\domain$ might have the sort of shape illustrated in Section~\ref{setting:domain}. As $\domain$ stretches out toward infinity, it has to part around the zeros of $p$, keeping well away from every zero except the one at $\zeta = 0$.
\begin{conditions}
\condconst{diag$_0$}{\lambda_\Delta}\label{cond:diag-basic} We can find a constant $\lambda_\Delta$ with
\[ \big| \hardker(a, a') \big| \lesssim e^{\lambda_\Delta|\zeta(a)-\zeta(a')|} \]
over all $a$ in $\domain$ outside some neighborhood of $\zeta = 0$, and all $a' \in \domain$.
%%outside a neighborhood of $\zeta(a) = 0$ in $\domain^2$.
\end{conditions}
\subsubsection{The perturbed operator}\label{setting:perturbed}

@@

\subsection{Main results}\label{sec:results}
We want to show that when $\volterra$ is a regular singular Volterra operator of the kind described in Section~\ref{setting:perturbed}, the equation $f = \volterra f$ has a unique solution of a certain form. For the prototypical operator $\hardpart$ described in Section~\ref{setting:basic}, this solution can be written explicitly.
\begin{thm}\label{thm:basic_volterra}
Suppose $\hardpart$ satisfies {\em Condition \eqref{cond:sing}}. Then the equation
\[ f = \hardpart f \]
has the {\em prototype solution}
\begin{equation}\label{eqn:test_solution}
\solproto(a)=\frac{p(b)}{p(a)} \exp\left(-\int_{b}^{a}\frac{q}{p}\;d\zeta\right).
\end{equation}
Changing the base point $b \in \domain$ just multiplies $f_0$ by a non-zero constant.
\end{thm}
We'll prove this result in Section~\ref{sec:construction}.
\color{RoyalBlue}
\begin{rmk}
If $q$ has a limit at $\zeta = 0$, the solution $f_0$ from Theorem~\ref{thm:basic_volterra} can be written as a constant mutiple of $\zeta^{\tau-1}$ plus a function which is $O(\zeta^\tau)$ at $\zeta = 0$.

\textcolor{SeaGreen}{$f_0 \in F\zeta^{\tau-1} + O(\zeta^\tau)$ near $\zeta = 0$ for some non-zero constant $F$.}
\end{rmk}
\color{black}
The solution $\solproto$ from Theorem~\ref{thm:basic_volterra} has, at worst, a mild power-law singularity at $\zeta = 0$. With stronger constraints on $\hardpart$, we can also ensure that $\solproto$ grows at most exponentially as $|\zeta| \to \infty$. The function spaces defined in Section~\ref{fn-spaces} express both of these regularity properties.
\begin{thm}\label{thm:proto-growth}
Suppose $\hardpart$ satisfies {\em Condition~\eqref{cond:sing}}. Then, on a small enough neighborhood of $\zeta = 0$, we have $|\solproto| \lesssim |\zeta|^{\tau-1}$.

@@

@@

%%To prove Theorem~\ref{thm:general_volterra}, we will rewrite the homogeneous equation we want to solve as an inhomogeneous one, with $\softpart f_0$ as the inhomogeneous term.

@@

\subsubsection{Construction}\label{sec:construction}
\begin{proof}[Proof of Theorem~\ref{thm:basic_volterra}]
Let
\[ \psi_0(a) = \exp\left(-\int_{b}^{a}\frac{q}{p}\;d\zeta\right), \]
so $\solproto = \frac{p(b)}{p} \psi_0$. Observing that $d\psi_0 = -\frac{q}{p} \psi_0\;d\zeta$, we can write $\hardpart \solproto$ in terms of an integral of $d\psi_0$. For each $a \in \Omega$, we have
\begin{align*}
\big[\hardpart \solproto\big](a) &= - \int_{\zeta=0}^{a} \frac{q}{p(a)}\,\solproto\;d\zeta \\
&= -\int_{\zeta=0}^{a} \frac{q}{p(a)}\,\frac{p(b)}{p}\psi_0\;d\zeta \\
&= - \frac{p(b)}{p(a)}  \int_{\zeta=0}^{a} \frac{q}{p} \psi_0\;d\zeta\\
&= \frac{p(b)}{p(a)} \int_{\zeta=0}^{a}\;d\psi_0.
\end{align*}

To evaluate this improper integral using the fundamental theorem of calculus, we need to find $\lim_{a \to 0} \psi_0(a)$. This limit turns out to be $0$, as shown by the following argument. By Condition~\eqref{cond:sing}, we can find a radius $\delta>0$ and a constant $C$ such that
\begin{equation}\label{eqn:limit_h1}
\left|\frac{q(a')}{p(a)} - \frac{\tau}{\zeta(a)}\right| < C
\end{equation}
whenever $|\zeta(a)| < \delta$ and $|\zeta(a')| < \delta$. Recalling that a change of base point just multiplies $\solproto$ by a non-zero constant, choose the base point $b \in \Omega$ so that $|\zeta(b)| = \delta/2$. \textcolor{orange}{[...]}
\begin{align*}
-\int_b^a \frac{q}{p}\;d\zeta & = \int_b^a \frac{\tau}{\zeta}\;d\zeta - \int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta \\
& = \tau \log\left(\frac{\zeta(a)}{\zeta(b)}\right) - \int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta
\end{align*}
Exponentiating both sides, we see that
\[ \psi_0(a) = \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau \exp\left[-\int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta\right]. \]
The exponential factor is bounded between $\exp\big({-\tfrac{3}{2}C\delta}\big)$ and $\exp\big(\tfrac{3}{2}C\delta\big)$ for all points $a$ in our chosen neighborhood of $\zeta = 0$. We deduce that $\hardpart  f_0=f_0$. 
\end{proof}

@@

@@

>>>>

\item \textcolor{gray}{Replace $\epsilon$ with $\gamma$.}

@@

Our definition of a regular singularity implies that as $|\zeta|$ goes to zero, $|p|$ also goes to zero, while $|q|$ settles into a compact interval that doesn't include zero.
\begin{verify}
[\textit{Proof.} Fix any $a'$ in the neighborhood where the difference above is bounded. By confining $a$ to small enough neighborhoods of $\zeta = 0$, we can make $|\tau/\zeta(a)|$ arbitrarily large, forcing $|q(a')/p(a)|$ to become arbitrarily large to satisfy the definition. Since $a'$ is fixed, the only way for this ratio to get arbitrarily large is for $|p(a)|$ to get arbitrarily small.]
\end{verify}
For some of our results, we'll also need to control $|p|$ and $|q|$ in various ways as $|\zeta|$ goes to infinity.
\begin{conditions}
\condconst{slow}{\lambda_0}\label{cond:slow} We can find constants $\lambda_p$ and $\lambda_\text{ratio}$ with
\begin{align*}
\left|\frac{1}{p}\right| & \lesssim e^{\lambda_p |\zeta|} &
\left|\frac{q}{p}\right| & \le \lambda_\text{ratio}
\end{align*}
outside some neighborhood of $\zeta = 0$ in $\Omega$. Note that this only constrains the ratio of $p$ and $q$ when they're evaluated at the same point. The constants will only enter our reasoning through their sum, $\lambda_0 = \lambda_p + \lambda_\text{ratio}$.
\end{conditions}
This condition explains why $\domain$ might have the sort of shape illustrated in Section~\ref{setting:domain}. As $\domain$ stretches out toward infinity, it has to part around the zeros of $p$, keeping well away from every zero except the one at $\zeta = 0$.
\begin{conditions}
\condconst{diag$_0$}{\lambda_\Delta}\label{cond:diag-basic} We can find a constant $\lambda_\Delta$ with
\[ \big| \hardker(a, a') \big| \lesssim e^{\lambda_\Delta|\zeta(a)-\zeta(a')|} \]
over all $a$ in $\domain$ outside some neighborhood of $\zeta = 0$, and all $a' \in \domain$.
%%outside a neighborhood of $\zeta(a) = 0$ in $\domain^2$.
\end{conditions}
We'll occasionally mention an optional condition on $p$ that allows us to state our main results more explicitly.
\begin{conditions}
\condconst{reg-p}{B, \epsilon}\label{cond:reg-p}
When $p$ is restricted to a small enough neighborhood of $\zeta = 0$, it falls into the class
\[ B\zeta + O\big(|\zeta|^{1 + \textcolor{orange}{\epsilon}}\big) \]
for some non-zero constant $B$ and some $\textcolor{orange}{\epsilon} > 0$.
\end{conditions}
\subsubsection{The perturbed operator}\label{setting:perturbed}

@@

\subsection{Main results}\label{sec:results}
We want to show that when $\volterra$ is a regular singular Volterra operator of the kind described in Section~\ref{setting:perturbed}, the equation $f = \volterra f$ has a unique solution of a certain form. For the prototypical operator $\hardpart$ described in Section~\ref{setting:basic}, this solution can be written explicitly.
\begin{thm}\label{thm:basic_volterra}
Suppose $\hardpart$ satisfies {\em Condition \eqref{cond:sing}}. Then the equation
\[ f = \hardpart f \]
has the {\em prototype solution}
\begin{equation}\label{eqn:test_solution}
\solproto(a)=\frac{p(b)}{p(a)} \exp\left(-\int_{b}^{a}\frac{q}{p}\;d\zeta\right).
\end{equation}
Changing the base point $b \in \domain$ just multiplies $f_0$ by a non-zero constant.
\end{thm}
We'll prove this result in Section~\ref{sec:construction}.

The solution $\solproto$ from Theorem~\ref{thm:basic_volterra} has, at worst, a mild power-law singularity at $\zeta = 0$. With stronger constraints on $\hardpart$, we can also ensure that $\solproto$ grows at most exponentially as $|\zeta| \to \infty$. The function spaces defined in Section~\ref{fn-spaces} express both of these regularity properties.
\begin{thm}\label{thm:proto-growth}
Suppose $\hardpart$ satisfies {\em Condition~\eqref{cond:sing}}. Then, on a small enough neighborhood of $\zeta = 0$, we have $|\solproto| \lesssim |\zeta|^{\tau-1}$.

@@

\begin{rmk}
Under stronger conditions on $p$, we can get a better estimate of the prototype solution near $\zeta = 0$, as described in Proposition~\ref{prop:better-proto-estimate} from Section~\ref{sec:asymptotics} for details.

%%Suppose $p$ falls into the class $B\zeta + O\big(|\zeta|^2\big)$, for some constant $B$, when it's restricted to some neighborhood of $\zeta = 0$.

%%for some constant $B$, we can strengthen the conclusion of Theorem~\ref{thm:proto-growth}. Under {\em Condition~\eqref{cond:sing}}, we can show that $\solproto \in B\zeta^{\tau-1} + O(|\zeta|^\tau)$.

%%If $q$ has a limit at $\zeta = 0$, the solution $f_0$ from Theorem~\ref{thm:basic_volterra} can be written as a constant multiple of $\zeta^{\tau-1}$ plus a function which is $O(\zeta^\tau)$ at $\zeta = 0$.
\end{rmk}
\color{black}

@@

\begin{rmk}
Under stronger conditions on $p$, Theorem~\ref{thm:general_volterra} can be restated to provide a unique solution in the affine subspace
\[ \zeta^{\tau-1} + \singexpalg{\tau-1+\gamma}(\Omega). \]
as described in Proposition~\ref{prop:alt-general_volterra} from Section~\ref{sec:existence and uniqueness}.
\end{rmk}
%%To prove Theorem~\ref{thm:general_volterra}, we will rewrite the homogeneous equation we want to solve as an inhomogeneous one, with $\softpart f_0$ as the inhomogeneous term.

@@

\subsubsection{Construction}\label{sec:construction}
\begin{proof}[Proof of Theorem~\ref{thm:basic_volterra}]
Let
\[ \psi_0(a) = \exp\left(-\int_{b}^{a}\frac{q}{p}\;d\zeta\right), \]
so $\solproto = \frac{p(b)}{p} \psi_0$. Observing that $d\psi_0 = -\frac{q}{p} \psi_0\;d\zeta$, we can write $\hardpart \solproto$ in terms of an integral of $d\psi_0$. For each $a \in \Omega$, we have
\begin{align*}
\big[\hardpart \solproto\big](a) &= - \int_{\zeta=0}^{a} \frac{q}{p(a)}\,\solproto\;d\zeta \\
& = -\int_{\zeta=0}^{a} \frac{q}{p(a)}\,\frac{p(b)}{p}\psi_0\;d\zeta \\
& = - \frac{p(b)}{p(a)}  \int_{\zeta=0}^{a} \frac{q}{p} \psi_0\;d\zeta\\
& = \frac{p(b)}{p(a)} \int_{\zeta=0}^{a}\;d\psi_0 \\
& = \frac{p(b)}{p(a)} \left[ \psi_0(a) - \lim_{\zeta \to 0} \psi_0 \right] \\
& = \solproto(a) - \frac{p(b)}{p(a)} \lim_{\zeta \to 0} \psi_0.
\end{align*}
Now, to prove that $\hardpart \solproto = \solproto$, we just need to show that $\lim_{\zeta \to 0} \psi_0 = 0$.

By Condition~\eqref{cond:sing}, we can find a radius $\delta>0$ and a constant $C$ such that
\begin{equation}\label{eqn:limit_h1}
\left|\frac{q(a')}{p(a)} - \frac{\tau}{\zeta(a)}\right| < C
\end{equation}
whenever $|\zeta(a)| < \delta$ and $|\zeta(a')| < \delta$. We can take advantage of this bound by rewriting the integral in the definition of $\psi_0$:
\begin{align*}
-\int_b^a \frac{q}{p}\;d\zeta & = \int_b^a \frac{\tau}{\zeta}\;d\zeta - \int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta \\
& = \tau \log\left(\frac{\zeta(a)}{\zeta(b)}\right) - \int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta
\end{align*}
Exponentiating both sides, we see that
\[ \psi_0(a) = \left(\frac{\zeta(a)}{\zeta(b)}\right)^\tau \exp\left[-\int_b^a \left( \frac{q}{p} + \frac{\tau}{\zeta} \right)\;d\zeta\right]. \]
Recalling that a change of base point just multiplies $\solproto$ by a non-zero constant, choose the base point $b \in \Omega$ so that $|\zeta(b)| = \delta/2$. The exponential factor in the formula above is then bounded between $\exp\big({-\tfrac{3}{2}C\delta}\big)$ and $\exp\big(\tfrac{3}{2}C\delta\big)$ whenever $|\zeta(a)| < \delta$. Since $\tau$ is positive, this is enough to show that $\lim_{\zeta \to 0} \psi_0 = 0$. It follows, as discussed above, that $\hardpart \solproto = \solproto$. 
\end{proof}

@@

\begin{prop}\label{prop:better-proto-estimate}
Suppose $p$ falls into the class $B\zeta + O\big(|\zeta|^{1 + \textcolor{orange}{\epsilon}}\big)$, for some constants $B$ and $\textcolor{orange}{\epsilon}$, when it's restricted to some neighborhood of $\zeta = 0$. In addition, suppose $\hardpart$ satisfies Condition~\eqref{cond:sing}. Then $\solproto \in B\zeta^{\tau-1} + O(|\zeta|^{\tau-1+\textcolor{orange}{\epsilon}})$.
\end{prop}
\begin{proof}

@@

\begin{prop}\label{prop:alt-general_volterra}
    Suppose $p$ satisfies Condition \eqref{cond:reg-p}, and $\volterra$ satisfies assumptions of Theorem \ref{thm:general_volterra}. Then equation 
    \[f=\volterra f\]
    has a unique solution $f$ in the affine subspace
    \[ B \zeta^{\tau-1}+\singexpalg{\tau-1+\gamma'}(\domain)\]
    of the space $\singexpalg{\tau-1}(\domain)$, where $\gamma'=\min\{\epsilon,\gamma\}$. 
\end{prop}
